The mass distribution of companions to low-mass white dwarfs

### 2014-10-24 documents/letter/ms.tex
text changes

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars
is (in general) not possible because of the unknown orbital plane inclination angles. 
Even when the mass of the visible star can be measured, only a lower limit can be placed
on the mass of the unseen companion. However, since these inclination angles should be isotropically
distributed, for a large enough, unbiased sample, the 
companion mass distribution can be deconvolved from the distributions of observables.
In this work, we construct a hierarchical probabilistic model to infer properties of unseen 
companion stars to low mass white dwarfs (LMWD) given orbital period and projected radial velocity measurements of the primary. We use a mixture of 
two Gaussians to model the true companion mass distribution to represent 
populations of white dwarfs (WD) and neutron stars (NS). 
Our model successfully recovers the initial parameters of several mock data sets.
We then apply our model to 55 of the 61 white dwarfs in 
the extremely low mass (ELM) WD survey. We find maximum a posteriori parameters for
the WD companion population of $\mu_{\rm WD} = 0.71~\Msun$ and $\sigma_{\rm WD} = 0.26~\Msun$ 
with a NS fraction consistent with zero, $f_{\rm NS} = 0\%$. However, the marginal posterior 
distribution over NS fraction has a significant tail up to $f_{\rm NS} \approx 16\%$.
We make samples from the posterior distribution publicly available so that future observational
efforts may compute the NS probability for newly discovered low mass WDs.
\end{abstract}

@@

The ELM sample is now large enough that the companion mass distribution, and importantly the NS fraction can be statistically constrained. In this work, we develop a probabilistic model to infer parameters of an assumed form for the mass distribution of companions to ELM WDs. Our method is similar to that employed by \citet{ozel12} and \citet{kiziltan13} to describe the mass distribution of NSs in binaries using post-Keplerian parameters. We focus on several specific astrophysical questions: 1.\ Can the population as a whole be modeled using a simple description of the companion masses? 2.\ How does the companion mass distribution compare to predictions from population synthesis simulations?  3.\ What is the rate of NS-LMWD binaries implied by our model? 4.\ What are the resulting distributions of NS probabilities for individual systems in the ELM sample?

@@

where the product is over each of the $j$ systems.  This marginal likelihood involves integrals over the latent variables $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_{-\infty}^\infty dM_2 \int_0^{\pi/2} di  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{The fractional uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$.} and assume the inclination angles are isotropically distributed
\begin{equation}
	p(\mf \given M_1, M_2, i) = \delta \left[\mf - f(M_1, M_2, i) \right]
\end{equation}
where
\begin{equation}
	f(M_1, M_2, i) = \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}
\end{equation}
and
\begin{equation}
	p(i) = \sin i.
\end{equation}
For now, we do not specify a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is
\begin{align}
%    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di \nonumber \\
%    & \qquad {} \times \delta \left[\mf - f(M_1, M_2, i)\right]~\sin i\\
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta})  \nonumber \\
    & \qquad {} \times \int_0^{\pi/2} di ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \mf - \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i.
\end{equation}
The inner integral (over $i$) has the form
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are 
\begin{align}
%	\sin i^* &= \alpha\\
%	\alpha &= \frac{(\mf(M_1+M_2)^2)^{1/3}}{M_2}\\
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
%	\frac{\partial g}{\partial i} &= \frac{3M_2^3}{(M_1+M_2)^2}\sin^2 i \cos i\\
%	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
We may rewrite the marginal likelihood as
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
%	&= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\alpha\left[\frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}\right]^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}

@@

We must now choose a functional form for the companion mass distribution,  $p(M_2\given \bs{\theta})$. For all experiments -- including that with the real data -- we use a two component Gaussian mixture model to fit the data. In tests below, we generate test data with different mixture component forms. We truncate the distributions using physically motivated bounds so that the white dwarf component is restricted to the range $M_2\in (0.2,\wdupper)~\Msun$ and the neutron star component is restricted to the range $M_2\in (1.3,2)~\Msun$,
\begin{align}
%	h(M_2, \mf, M_1) &= \frac{(M_1+M_2)^{4/3}}{3\mf^{1/3}M_2\sqrt{M_2^2 - \left[ \mf(M_1+M_2)^2 \right]^{2/3}}}\\
%	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~\left[ (1-f_{\rm NS})p_{\rm WD} + f_{\rm NS}p_{\rm NS} \right]~h(M_2, \mf, M_1)\\
	p(M_2 \given \bs{\theta}) &= \left[ (1-f_{\rm NS})~p_{\rm WD} + f_{\rm NS}~p_{\rm NS} \right] 
\end{align}
where 
\begin{align}
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD}); ~0.2 < \frac{M_2}{\Msun} < \wdupper \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS}); ~1.3 < \frac{M_2}{\Msun} < 2.
\end{align}
and $\mathcal{N}$ is the (truncated, but properly normalized) normal distribution with mean $\mu$ and variance $\sigma^2$ and the distributions are limited to the ranges specified. In this work, we fix the mean and variance of the neutron star distribution to
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun\\
	\sigma^2_{\rm NS} &= (0.05~\Msun)^2
\end{align}
based on indications that NSs in certain binaries may be somewhat more massive than the canonical NS mass of 1.35 \Msun~ \citep{kiziltan13,smedley14}. The probability of any particular system having a NS companion $P_{\rm NS}$ is then:
\begin{equation}
P_{\rm NS} = \int_{0}^{\infty} dM_2~ f_{\rm NS}~ p_{\rm NS}~ h(M_2, \mf, M_1) \label{eq:P_NS}
\end{equation}

Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$. We choose uninformative priors on these parameters: for the mean of the white dwarf component, we use a uniform distribution from $0.2-1.4~\Msun$; we use a (scale-invariant) logarithmic prior for the variance of the white dwarf component over the range $0.05-2~\Msun$; finally, we use a uniform prior over the dimensionless neutron star fraction from $0-1$. The model parameters are summarized in Table~\ref{tbl:parameters}.

@@

We use an ensemble Markov Chain Monte Carlo algorithm \citep{goodman10} to draw samples from the posterior distribution, $p(\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS} \given \bs{m}_f, \bs{M}_1)$. The algorithm is implemented in the \texttt{Python} programming language \citep[{\tt emcee};][]{foremanmackey13} and uses an ensemble of individual ``walkers'' to naturally adapt to the geometry of the parameter-space being explored. We run the walkers for an initial, burn-in period of 500 steps starting from randomly drawn initial conditions (sampled from the priors in Table~\ref{tbl:parameters}). We then re-initialize the walkers from their positions at the end of this run and run again for 1000 steps. We throw out the burn-in samples to eliminate any effects from our choice of initial conditions. The autocorrelation times for each parameter confirm they have converged.

@@

As a first test, we generate $M_2$ by drawing from a single, truncated Gaussian with parameters given in Table~\ref{tbl:parameters}. This test sample contains no NSs. In the top row of Figure~\ref{fig:ELM_post}, the leftmost panel shows that the maximum a posteriori (MAP) $M_2$ distribution (black line) qualitatively matches the input distribution (grey histogram). The second panel shows samples from the posterior distribution and contours that contain 68\% and 95\% of the samples. The input values lie cleanly within the inner contour in both projections of the posterior shown; the model preference toward higher masses and smaller standard deviations is due to statistical noise from our randomly generated mass distribution. The third panel shows that although the posterior $f_{\rm NS}$ distribution is consistent with 0\%, there is a tail up to $\sim$10\%. 
%Nevertheless, our model recovers the input parameters for the white dwarf companion population. 
For any set of model parameters, Equation~\ref{eq:P_NS} gives the probability of an individual system hosting a NS. using samples from our posteriors, we can determine the distribution of $P_{\rm NS}$ for each system. The fourth panel lists individual systems, ordered by $\mf$, and shows the distributions of $P_{\rm NS}$ for each. Most systems have their entire distribution of $P_{\rm NS}$ values less than $\sim$5\%.

\subsection{Test 2: Two Gaussians (WD + NS)} \label{sec:exp2}
For our next test, we use the same Gaussian distribution to generate companion masses for the WD component as the previous example, but we add an additional NS component with $f_{\rm NS} = 10\%$ and parameters described in Section~\ref{sec:experiments}. The middle row of Figure~\ref{fig:ELM_post} shows that our model again recovers the input values for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. As in Test 1, the model preference for higher $\mu_{\rm WD}$ and lower $\sigma_{\rm WD}$ is due to our randomly generated mass distribution. Importantly, the third panel shows that it also recovers $f_{\rm NS}$, however the posterior shows a substantial tail toward higher probabilities. The fourth panel indicates that some systems have near unity $P_{\rm NS}$ values; our model correctly identifies these as true NSs. However this accounts for less than half of the systems with NSs in the simulated sample; the other half have low $\mf$'s due to low inclination angles. Such systems are impossible to statistically differentiate them from those with WD companions.

\subsection{Test 3: Uniform (WD) + Gaussian (NS)} \label{sec:exp3}
As a final test, we generate companion masses for the WD component by sampling from a uniform distribution over the range $(0.2,1.2)~\Msun$ while keeping $f_{\rm NS}=$10\%. The bottom row of Figure~\ref{fig:ELM_post} shows the results of this test. The posterior distribution in the second panel shows that the model does not converge on any one solution for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. The preference toward larger $\sigma_{\rm WD}$ is expected as it flattens the distribution to match it with the input uniform distribution. The third panel shows that the posterior distribution is spread out in $\sigma_{\rm WD}$. Interestingly, despite having a non-Gaussian input distribution for $M_2$, our model still recovers $f_{\rm NS}$ approximately as accurately as in Test 2. The fourth panel again shows that, compared to Test 2, our model is not as effective at identifying which LMWDs host NS companions.

@@

The results from applying our model to the ELM sample are shown in Figure \ref{fig:ELM_post}. The MAP model has parameters $\mu_{\rm WD} = 0.71~\Msun$, $\sigma_{\rm WD} = 0.26~\Msun$, and $f_{\rm NS} = 0\%$. The marginal posterior over $\mu_{\rm WD}$ and $\sigma_{\rm WD}$ has a tail towards larger variances, which could indicate that the true WD distribution is not exactly Gaussian. It is interesting that the best fit Gaussian for the companions to the ELM WD sample is similar to that of the population of single DA WDs in SDSS with a mean of 0.6 $\Msun$ \citep{kleinman13}. Our distribution is significantly wider ({\bf $\sigma \approx 0.26~\Msun$} compared with $\sigma \approx 0.1~\Msun$), possibly due to mass transfer affecting the mass distribution of the unseen primary WDs as well.

@@

The third panel in Figure~\ref{fig:ELM_post} shows a NS fraction strongly peaked toward 0. However, there is a significant tail toward higher NS probabilities. Our model indicates $f_{\rm NS} <16\%$ at the 68\% confidence level, in agreement with independent constraints from \citet{vLeeuwen07} ($f_{\rm NS}<18\pm5$\%), based on the non-detection of radio emission from eight LMWD targets, and from \citet{agueros09b} ($f_{\rm NS}<10\substack{+4 \\ -2}~\%$), based on non-detections of radio emission from an expanded sample of 15 LMWD targets in the ELM WD sample. Because it may be radio quiet or beamed away from us, a NS companion to these LMWDs may not be detectable \citep{vLeeuwen07}. All NSs give off blackbody radiation peaked in X-ray wavelengths, so X-ray follow-up provides a more stringent test on the potential for a NS companion \citep{agueros09a}. Previous searches for LMWDs with X-ray counterparts have all been non-detections \citep{ELMII,ELMIV,kilic13,kilic14}. However, we show here that all of those sources have a low probability of hosting a NS companion. If the two systems with the highest $P_{\rm NS}$, SDSS J081133.6$+$022556.8 and SDSS J174140.5$+$652638.7, also lack X-ray counterparts, then even more stringent constraints can be placed on $f_{\rm NS}$.

@@

We have developed a statistical model to infer the companion mass distribution for a sample of single-lined, spectroscopic binaries. This model can be applied to any such sample with measured $M_1$. When applied to three separate test cases with WD and NS companion masses drawn from different distributions, our model recovers the correct input parameters; even if the companion mass distribution is not drawn from a Gaussian distribution, our model still infers the input NS fraction within a few percent.

We then applied our model to the set of WDs from the ELM WD survey. Our model indicates that the fraction of ELM WDs with NS companions is consistent with 0\%, but could be up to $\sim$16\% (within 1-$\sigma$). We identify SDSS J081133.6$+$022556.8 and SDSS J174140.5$+$652638.7 as having the highest probabilities of hosting a NS companion and suggest that X-ray and radio follow-up is warranted. Our model further suggests that the majority of ELM WDs have CO-core WD companions. This is in contrast to predictions from population synthesis models which predict the dominant companion population should be He-core WDs \citep{toonen12}.

>>>>

\begin{abstract}
Measuring the masses of companions to single-line spectroscopic binary stars
is (in general) not possible because of the unknown orbital plane inclination angles. 
Even when the mass of the visible star can be measured, only a lower limit can be placed
on the mass of the unseen companion. However, since these inclination angles should be isotropically
distributed, for a large enough, unbiased sample, the 
companion mass distribution can be deconvolved from the distributions of observables.
In this work, we construct a hierarchical probabilistic model to infer properties of unseen 
companion stars given observations of the orbital period and projected radial velocity of the primary star. 
We then apply this model to a sample of low mass white dwarfs (LMWD).
We use a mixture of 
two Gaussians to model the true companion mass distribution to represent 
populations of white dwarfs (WD) and neutron stars (NS). 
Our model successfully recovers the initial parameters of several mock data sets.
We then apply our model to 55 of the 61 white dwarfs in 
the extremely low mass (ELM) WD survey. We find maximum a posteriori parameters for
the WD companion population of $\mu_{\rm WD} = 0.71~\Msun$ and $\sigma_{\rm WD} = 0.26~\Msun$ 
with a NS fraction consistent with zero, $f_{\rm NS} = 0\%$. However, the marginal posterior 
distribution over NS fraction has a significant tail up to $f_{\rm NS} \approx 16\%$.
We make samples from the posterior distribution publicly available so that future observational
efforts may compute the NS probability for newly discovered low mass WDs.
\end{abstract}

@@

The ELM sample is now large enough that the companion mass distribution, and importantly the NS fraction, can be statistically constrained. In this work, we develop a probabilistic model to infer parameters of an assumed form for the mass distribution of companions to ELM WDs. Our method is similar to that employed by \citet{ozel12} and \citet{kiziltan13} to describe the mass distribution of NSs in binaries using post-Keplerian parameters. We focus on several specific astrophysical questions: 1.\ Can the population as a whole be modeled using a simple description of the companion masses? 2.\ How does the companion mass distribution compare to predictions from population synthesis simulations?  3.\ What is the rate of NS-LMWD binaries implied by our model? 4.\ What are the resulting distributions of NS probabilities for individual systems in the ELM sample?

@@

where the product is over each of the $j$ systems.  This marginal likelihood involves integrals over the unobserved quantities $i$ and $M_2$,
\begin{align}
    p(\mf \given \bs{\theta}) &= \int_{-\infty}^\infty dM_2 \int_0^{\pi/2} di  \nonumber \\
      & \qquad {} \times p(\mf \given M_1, M_2, i)~p(M_2 \given \bs{\theta})~p(i).
\end{align}
We neglect observational uncertainties in $\mf$ and $M_1$,\footnote{The fractional uncertainties in these quantities are small, $\sigma_x / x \sim 0.05-0.1$.} and assume the inclination angles are isotropically distributed
\begin{equation}
	p(\mf \given M_1, M_2, i) = \delta \left[\mf - f(M_1, M_2, i) \right]
\end{equation}
where
\begin{equation}
	f(M_1, M_2, i) = \frac{(M_2 \sin i)^3}{(M_1 + M_2)^2}
\end{equation}
and
\begin{equation}
	p(i) = \sin i.
\end{equation}
For now, we do not specify a parametric form for the companion mass distribution, $p(M_2 \given \bs{\theta})$. With the above assumptions, the marginal likelihood integral is
\begin{align}
%    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta}) \int_0^{\pi/2} di \nonumber \\
%    & \qquad {} \times \delta \left[\mf - f(M_1, M_2, i)\right]~\sin i\\
    p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2 ~p(M_2 \given \bs{\theta})  \nonumber \\
    & \qquad {} \times \int_0^{\pi/2} di ~\sin i ~ \delta \left[g(M_1,M_2,i) \right]\label{eq:delta}
\end{align}
where
\begin{equation}
	g(M_1,M_2,i) = \mf - \frac{M_2^3}{(M_1+M_2)^2}\sin^3 i.
\end{equation}
The inner integral (over $i$) has the form
\begin{equation}
    \int dx~F(x)~\delta \left[ G(x) \right] = \sum_j \frac{F(x^*_j)}{|G'(x^*_j)|}
\end{equation}
where the sum is over the roots, $x^*_j$, of the function $G(x)$. The root, $i^*$, and derivative of the argument of the delta function in \eqn\ref{eq:delta} are 
\begin{align}
%	\sin i^* &= \alpha\\
%	\alpha &= \frac{(\mf(M_1+M_2)^2)^{1/3}}{M_2}\\
	\sin i^* &= \frac{ \left[\mf(M_1+M_2)^2 \right]^{1/3}}{M_2}\\
%	\frac{\partial g}{\partial i} &= \frac{3M_2^3}{(M_1+M_2)^2}\sin^2 i \cos i\\
%	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}
	\frac{\partial g}{\partial i}\bigg\rvert_{i^*} &= \frac{3M_2^3}{(M_1+M_2)^2} \sin^2 i^* \sqrt{1 - \sin^2 i^*}
\end{align}
We may rewrite the marginal likelihood as
\begin{align}
	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\sin i^* \left(\frac{\partial g}{\partial i}\bigg\rvert_{i^*}\right)^{-1}\\
%	&= \int_{0}^\infty dM_2~p(M_2 \given \bs{\theta})~\alpha\left[\frac{3M_2^3}{(M_1+M_2)^2} \alpha^2 \sqrt{1 - \alpha^2}\right]^{-1}\\
	&= \int_{M_{2,{\rm min}}}^\infty dM_2~p(M_2 \given \bs{\theta})~h(M_2, \mf, M_1) \label{eq:fullm2}
\end{align}

@@

We must now choose a functional form for the companion mass distribution,  $p(M_2\given \bs{\theta})$. For all experiments -- including that with the real data -- we use a two component Gaussian mixture model to fit the data. In tests below, we generate test data with different mixture component forms. We truncate the distributions using physically motivated bounds so that the white dwarf component is restricted to the range $M_2\in (0.2,\wdupper)~\Msun$ and the neutron star component is restricted to the range $M_2\in (1.3,2)~\Msun$,
\begin{align}
%	h(M_2, \mf, M_1) &= \frac{(M_1+M_2)^{4/3}}{3\mf^{1/3}M_2\sqrt{M_2^2 - \left[ \mf(M_1+M_2)^2 \right]^{2/3}}}\\
%	p(\mf \given \bs{\theta}) &= \int_{0}^\infty dM_2~\left[ (1-f_{\rm NS})p_{\rm WD} + f_{\rm NS}p_{\rm NS} \right]~h(M_2, \mf, M_1)\\
	p(M_2 \given \bs{\theta}) &= \left[ (1-f_{\rm NS})~p_{\rm WD} + f_{\rm NS}~p_{\rm NS} \right] 
\end{align}
where 
\begin{align}
	p_{\rm WD} &= \mathcal{N}(M_2 \given \mu_{\rm WD}, \sigma^2_{\rm WD}); ~0.2 < \frac{M_2}{\Msun} < \wdupper \\
	p_{\rm NS} &= \mathcal{N}(M_2 \given \mu_{\rm NS}, \sigma^2_{\rm NS}); ~1.3 < \frac{M_2}{\Msun} < 2.
\end{align}
and $\mathcal{N}$ is the (truncated, but properly normalized) normal distribution with mean $\mu$ and variance $\sigma^2$ and the distributions are limited to the ranges specified. In this work, we fix the mean and variance of the neutron star distribution to
\begin{align}
	\mu_{\rm NS} &= 1.4~\Msun\\
	\sigma^2_{\rm NS} &= (0.05~\Msun)^2
\end{align}
based on indications that NSs in certain binaries may be somewhat more massive than the canonical NS mass of 1.35 \Msun~ \citep{kiziltan13,smedley14}. The probability of any particular system having a NS companion, $P_{\rm NS}$, can be computed for a given set of parameters for the companion mass distribution:
\begin{equation}
P_{\rm NS} = \frac{\int_{M_{2,{\rm min}}}^{\infty} dM_2~ f_{\rm NS}~ p_{\rm NS}~ h(M_2, \mf, M_1)}{\int_{M_{2,{\rm min}}}^{\infty} dM_2~ \left[ (1-f_{\rm NS})~p_{\rm WD} + f_{\rm NS}~p_{\rm NS} \right] ~ h(M_2, \mf, M_1)}\label{eq:P_NS}
\end{equation}

Our companion mass model parameters are then $\bs{\theta} = (\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS})$. For the mean of the white dwarf component, we use a uniform prior from $0.2-1.4~\Msun$; we use a logarithmic prior for the variance of the white dwarf component over the range $0.05-2~\Msun$; finally, we use a uniform prior over the dimensionless neutron star fraction from $0-1$. The model parameters are summarized in Table~\ref{tbl:parameters}.

@@

We use an ensemble Markov Chain Monte Carlo algorithm \citep{goodman10} to draw samples from the posterior distribution, $p(\mu_{\rm WD}, \sigma_{\rm WD}, f_{\rm NS} \given \bs{m}_f, \bs{M}_1)$. The algorithm is implemented in the \texttt{Python} programming language \citep[{\tt emcee};][]{foremanmackey13} and uses an ensemble of individual ``walkers'' to naturally adapt to the geometry of the parameter-space being explored. We run the walkers for an initial, burn-in period of 500 steps starting from randomly drawn initial conditions (sampled from the priors in Table~\ref{tbl:parameters}). We then re-initialize the walkers from their positions at the end of this run and run again for 1000 steps. We throw out the burn-in samples to eliminate any effects from our choice of initial conditions. The autocorrelation times for each parameter indicate that they have converged.

@@

As a first test, we generate $M_2$ by drawing from a single, truncated Gaussian with parameters given in Table~\ref{tbl:parameters}. This test sample contains no NSs. In the top row of Figure~\ref{fig:ELM_post}, the leftmost panel shows that the maximum a posteriori (MAP) $M_2$ distribution (black line) qualitatively matches the input distribution (grey histogram). The second panel shows samples from the posterior distribution and contours that contain 68\% and 95\% of the samples. The input values lie cleanly within the inner contour in both projections of the posterior shown; the model preference toward higher masses and smaller standard deviations is due to statistical noise from our randomly generated test sample. The third panel shows that although the posterior $f_{\rm NS}$ distribution is consistent with 0\%, there is a tail up to $\sim$10\%. 
%Nevertheless, our model recovers the input parameters for the white dwarf companion population. 
For any set of model parameters, Equation~\ref{eq:P_NS} gives the probability of an individual system hosting a NS. Using posterior samples, we can determine the distribution of $P_{\rm NS}$ for each system. The fourth panel lists individual systems, ordered by $\mf$, and shows the distributions of $P_{\rm NS}$ for each. For most systems, there is negligible probability above $P_{\rm NS}\sim 5\%$.

\subsection{Test 2: Two Gaussians (WD + NS)} \label{sec:exp2}
For our next test, we use the same Gaussian distribution to generate companion masses for the WD component as the previous example, but we add an additional NS component with $f_{\rm NS} = 10\%$ and parameters described in Section~\ref{sec:experiments}. The middle row of Figure~\ref{fig:ELM_post} shows that our model again recovers the input values for $\mu_{\rm WD}$ and $\sigma_{\rm WD}$. As in Test 1, the model preference for higher $\mu_{\rm WD}$ and lower $\sigma_{\rm WD}$ is due to our randomly generated mass distribution. Importantly, the third panel shows that it also recovers $f_{\rm NS}$, however the posterior shows a substantial tail toward higher probabilities. The fourth panel indicates that some systems have near unity $P_{\rm NS}$ values; our model correctly identifies these as true NSs. However this accounts for less than half of the systems with NSs in the simulated sample; the other half have low $\mf$'s due to low inclination angles. Such systems are impossible to statistically differentiate from those with WD companions.

\subsection{Test 3: Uniform (WD) + Gaussian (NS)} \label{sec:exp3}
As a final test, we generate companion masses for the WD component by sampling from a uniform distribution over the range $(0.2,1.2)~\Msun$, again with $f_{\rm NS}=$10\%. The bottom row of Figure~\ref{fig:ELM_post} shows the results of this test. The posterior distribution in the second panel shows that $\mu_{\rm WD}$ and $\sigma_{\rm WD}$ are not well constrained, as expected. The preference toward larger $\sigma_{\rm WD}$ is expected as it flattens the distribution to match it with the input uniform distribution. The third panel shows that the posterior distribution is spread out in $\sigma_{\rm WD}$. Interestingly, despite having a non-Gaussian input distribution for $M_2$, our model still recovers $f_{\rm NS}$ approximately as accurately as in Test 2. The fourth panel again shows that, compared to Test 2, our model is not as effective at identifying which LMWDs host NS companions.

@@

The results from applying our model to the ELM sample are shown in Figure \ref{fig:ELM_post}. The MAP model has parameters $\mu_{\rm WD} = 0.71~\Msun$, $\sigma_{\rm WD} = 0.26~\Msun$, and $f_{\rm NS} = 0\%$. The marginal posterior over $\mu_{\rm WD}$ and $\sigma_{\rm WD}$ has a tail towards larger variances, which could indicate that the true WD distribution may not be exactly Gaussian. It is interesting that the best fit Gaussian for the companions to the ELM WD sample is similar to that of the population of single DA WDs in SDSS with a mean of 0.6 $\Msun$ \citep{kleinman13}. Our distribution is significantly wider ({\bf $\sigma \approx 0.26~\Msun$} compared with $\sigma \approx 0.1~\Msun$), possibly due to mass transfer affecting the mass distribution of the unseen primary WDs as well.

@@

The third panel in Figure~\ref{fig:ELM_post} shows a NS fraction strongly peaked toward 0. However, there is a significant tail toward higher NS probabilities. Our model indicates $f_{\rm NS} <16\%$ at the 68\% confidence level, in agreement with independent constraints from \citet{vLeeuwen07} ($f_{\rm NS}<18\pm5$\%), based on the non-detection of radio emission from eight LMWD targets, and from \citet{agueros09b} ($f_{\rm NS}<10\substack{+4 \\ -2}~\%$), based on non-detections of radio emission from an expanded sample of 15 LMWD targets in the ELM WD sample. Because it may be radio quiet or beamed away from us, a NS companion to these LMWDs may not be detectable \citep{vLeeuwen07}. All NSs give off blackbody radiation peaked in X-ray wavelengths, so X-ray follow-up provides a more stringent test on the potential for a NS companion \citep{agueros09a}. Previous searches for LMWDs with X-ray counterparts have all been non-detections \citep{ELMII,ELMIV,kilic13,kilic14}. However, based on our model, all of those sources have a low probability of hosting a NS companion. 
%If the two systems with the highest $P_{\rm NS}$, SDSS J081133.6$+$022556.8 and SDSS J174140.5$+$652638.7, also lack X-ray counterparts, then even more stringent constraints can be placed on $f_{\rm NS}$.

@@

We have developed a statistical model to infer the companion mass distribution for a sample of single-line, spectroscopic binaries. This model can be applied to any such sample with measured $M_1$ and mass function. When applied to three separate test cases with WD and NS companion masses drawn from different distributions, our model recovers the input parameters; even if the companion mass distribution is not drawn from a Gaussian distribution, our model still infers the input NS fraction within a few percent.

We then applied our model to the set of WDs from the ELM WD survey. Our model indicates that the fraction of ELM WDs with NS companions is consistent with 0\%, but could be up to $\sim$16\% (within 1-$\sigma$). We identify SDSS J081133.6$+$022556.8 and SDSS J174140.5$+$652638.7 as having the highest median probabilities of hosting a NS companion and suggest that X-ray and radio follow-up is warranted. Our model further suggests that the majority of ELM WDs have CO-core WD companions. This is in contrast to predictions from population synthesis models which predict the dominant companion population should be He-core WDs \citep{toonen12}.
